---
title: "quallmer audit trail"
date: "2026-04-23"
format:
  html:
    toc: true
    toc-depth: 3
---

This audit trail documents the complete workflow following Lincoln and Guba's (1985)
framework for establishing trustworthiness in qualitative research.

## Trail summary

- **Number of runs:** 3
- **Complete:** Yes

### System information

- **quallmer version:** 0.3.0
- **ellmer version:** 0.4.0
- **R version:** 4.4.3

## Instrument development

Codebooks used in this analysis:

### Economic optimism (exercise)

**Instructions:**

> Rate how optimistic the speech is about the national economy
    over the coming years. Use an integer from 0 (very pessimistic) to 5 (very optimistic).

**Schema:**

```
<ellmer::TypeObject>
 @ description          : NULL
 @ required             : logi TRUE
 @ properties           :List of 2
 .. $ score      : <ellmer::TypeBasic>
 ..  ..@ description: chr "Economic optimism score 0--5"
 ..  ..@ required   : logi TRUE
 ..  ..@ type       : chr "integer"
 .. $ explanation: <ellmer::TypeBasic>
 ..  ..@ description: chr "Brief justification with concrete phrases from the text"
 ..  ..@ required   : logi TRUE
 ..  ..@ type       : chr "string"
 @ additional_properties: logi FALSE
```

## Process notes

Chronological record of all coding runs:

### 1. ex2_initial_coding

**Parent:** None (original run)
**Timestamp:** 2026-04-23 17:01:12
**Model:** openai/gpt-4o-mini
**Codebook:** Economic optimism (exercise)
**Units coded:** 5

**Call:**

```{r eval=FALSE}
qlm_code(x = my_texts, codebook = ex2_codebook, model = "openai/gpt-4o-mini", 
    name = "ex2_initial_coding")
```

### 2. ex2_replicate_same_params

**Parent:** ex2_initial_coding
**Timestamp:** 2026-04-23 17:01:16
**Model:** openai/gpt-4o-mini
**Codebook:** Economic optimism (exercise)
**Units coded:** 5

**Call:**

```{r eval=FALSE}
qlm_replicate(x = ex2_run1, name = "ex2_replicate_same_params")
```

### 3. ex2_replicate_high_temperature

**Parent:** ex2_initial_coding
**Timestamp:** 2026-04-23 17:01:21
**Model:** openai/gpt-4o-mini
**Codebook:** Economic optimism (exercise)
**Units coded:** 5

**Call:**

```{r eval=FALSE}
qlm_replicate(x = ex2_run1, params = params(temperature = 0.9), 
    name = "ex2_replicate_high_temperature")
```

## Raw data and coded results

The complete coded data for each run is stored in the RDS file.
Load with `readRDS()` to access:

```r
trail <- readRDS("path/to/trail.rds")
trail$runs$run_name$data
# Coded results for a specific run
```

**ex2_initial_coding:** 5 units, 2 variables
**ex2_replicate_same_params:** 5 units, 2 variables
**ex2_replicate_high_temperature:** 5 units, 2 variables

## Replication

This section provides instructions and code to replicate the analysis.

### Environment setup

Install required packages:

```r
# Install quallmer (if not already installed)
# install.packages("pak")
# pak::pak("quallmer/quallmer")

library(quallmer)
```

Configure API credentials for the models used:

```r
# OpenAI: Set OPENAI_API_KEY environment variable
# Sys.setenv(OPENAI_API_KEY = "your-api-key")
```

### Loading the trail

Load the saved trail to access codebooks, data, and metadata:

```r
trail <- readRDS("exercise2_workflow.rds")
```

### Replication code

The following code replicates each coding run using the stored parameters.

#### Replicate: ex2_initial_coding

```r
# Get codebook from trail
codebook_ex2_initial_coding <- trail$runs$ex2_initial_coding$codebook

# Load your input data (same structure as original)
# texts <- your_data$text_column

coded_ex2_initial_coding <- qlm_code(
  texts,
  codebook_ex2_initial_coding,
  model = "openai/gpt-4o-mini",
  name = "ex2_initial_coding"
)
```

#### Replicate: ex2_replicate_same_params

```r
# Get codebook from trail
codebook_ex2_replicate_same_params <- trail$runs$ex2_replicate_same_params$codebook

# Load your input data (same structure as original)
# texts <- your_data$text_column

coded_ex2_replicate_same_params <- qlm_code(
  texts,
  codebook_ex2_replicate_same_params,
  model = "openai/gpt-4o-mini",
  name = "ex2_replicate_same_params"
)
```

#### Replicate: ex2_replicate_high_temperature

```r
# Get codebook from trail
codebook_ex2_replicate_high_temperature <- trail$runs$ex2_replicate_high_temperature$codebook

# Load your input data (same structure as original)
# texts <- your_data$text_column

coded_ex2_replicate_high_temperature <- qlm_code(
  texts,
  codebook_ex2_replicate_high_temperature,
  model = "openai/gpt-4o-mini",
  name = "ex2_replicate_high_temperature"
)
```

### Note on reproducibility

LLM outputs are inherently stochastic. To improve reproducibility:

- Use `temperature = 0` for more deterministic outputs
- Set a random seed where supported by the provider
- Document the exact model version (models are updated over time)
- Compare results across multiple runs using `qlm_replicate()`

## Reference

Lincoln, Y. S., & Guba, E. G. (1985). *Naturalistic Inquiry*. Sage.

